---
title: "NHL_Shot_Analysis_Visualization"
author: "Daniel, Garrett, Ian"
date: "`r Sys.Date()`"
output: html_document
---

# NHL Shot Analysis Visualization 
### Load library
```{r}
library(ggpubr)
library(tidyverse)
library(forcats)
library(jpeg)
library(grid)
library(gganimate)
library(visdat)
library(eeptools)
library(glmnet) # Load glmnet
library(plotmo) # for plot_glmnet
library(rpart)                      # Popular decision tree algorithm
library(rattle)                 # Fancy tree plot
library(rpart.plot)             # Enhanced tree plots
library(RColorBrewer)               # Color selection for fancy tree plot
library(party)                  # Alternative decision tree algorithm
library(partykit)               # Convert rpart object to BinaryTree
library(caret)  
library(randomForest)
library(sparcl) # Sparse Clustering
library(cluster) # Load cluster
library(factoextra) # clustering algorithms & visualization
library(xgboost)
library(xgboostExplainer) # Load XGboost Explainer
library(pROC) # Load proc
```

## Original Data & Cleaning
Data Cleaning has been done in the Daniel Analysis.Rmd. Referring to Daniel Analysis.Rmd for details of cleaning.


### Load Clean Data
```{r}
NHL_Shot_Analysis_Data <- read_csv("NHL_Shot_Analysis_Data.csv", col_select = -1)
```
Data preview
```{r}
# How many rows and columns?
dim(NHL_Shot_Analysis_Data)
# What are the data types and what do they look like in general
summary(NHL_Shot_Analysis_Data)
# The names of all columns of the data
names(NHL_Shot_Analysis_Data)
# First 6 rows of the data
head(NHL_Shot_Analysis_Data)
# Last 6 rows of the data
tail(NHL_Shot_Analysis_Data)
# How many NAs are there?
sum(is.na(NHL_Shot_Analysis_Data))

```

NAs handling
```{r}
# Check the NAs in data set
sum(is.na(NHL_Shot_Analysis_Data))

# Remove all NAs in the data set
NHL_Shot_Analysis_Data <- na.omit(NHL_Shot_Analysis_Data)
```


## Data Visualization
Focus on the secondaryType of "shot" & "goal" in the event column. Calculate the frequency of each types of shots and their conversion rate.

```{r}
# Examine different types of shots
summary(as.factor(NHL_Shot_Analysis_Data$secondaryType))

# Create a table to summarize the frequency of different types of shots
shots.count.df <- NHL_Shot_Analysis_Data  %>%
  group_by(secondaryType) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

print(shots.count.df)

# Visualization of the frequency of different types of shots
options(scipen = 1) # No scientific notation

g1 <- count(NHL_Shot_Analysis_Data, secondaryType) %>%
  ggplot(aes(x = reorder(secondaryType, -n), y = n))+
  geom_col(aes(fill = as.factor(secondaryType))) +
  geom_label(aes(label = n), label.size = 0.1, vjust = -0.3)+
  scale_y_continuous(name = "Frequency", limits = c(0, 550000), breaks = c(0, 100000,200000,300000,400000,500000))+
  scale_x_discrete(name = "Types of Shot") +
  ggtitle(paste("Frequency of Types of Shot n =", sum(shots.count.df$count),sep = " "))

# Conversion Rate of Different Types of Shot
Shot.df <- NHL_Shot_Analysis_Data %>%
  filter(event == "Shot" & is.na(secondaryType) == F) %>%
  group_by(secondaryType) %>%
  summarise(count=n())

Goal.df <- NHL_Shot_Analysis_Data %>%
  filter(event == "Goal" & is.na(secondaryType) == F) %>%
  group_by(secondaryType) %>%
  summarise(count=n())

Shot_Conversion.df <- data.frame(Type_of_Shots = Shot.df$secondaryType, Conversion_Rate = round(Goal.df$count/Shot.df$count, digits = 2))

g2 <- ggplot(Shot_Conversion.df, aes(x = reorder(Type_of_Shots, -Conversion_Rate), y = Conversion_Rate))+
  geom_col(aes(fill = as.factor(Type_of_Shots)))+
  geom_label(aes(label = Conversion_Rate), label.size = 0.1, vjust = -0.3)+
  scale_y_continuous(name = "Conversion Rate", limits = c(0, 0.25), breaks = c(0, 0.05, 0.10, 0.15, 0.20, 0.25))+
  scale_x_discrete(name = "Types of Shot") +
  ggtitle("Conversion Rate of Different Types of Shot")

# Create a Side-by-Side Bar chart
g3 <- ggarrange(g1,g2,
                labels = c ("A", "B"),
                ncol = 1, nrow = 2,
                common.legend = T, legend = "right")

print(g3)
```


```{r}
# Add a graph of the rink
rink <- rasterGrob(readJPEG("ice_rink.jpg"),
                   width = unit(1.1,"npc"), height=unit(1.1, "npc"))
unique(NHL_Shot_Analysis_Data$secondaryType)

# Wrap-around
ggplot(NHL_Shot_Analysis_Data[NHL_Shot_Analysis_Data$secondaryType == "Wrap-around",],
       aes(x = x, y = y, group = event, color = event, size = event))+
  annotation_custom(rink, -100, 100, -40, 40) + 
  geom_point() +
  theme_bw() +
  theme(panel.grid.major = element_blank(), # Remove grid
        panel.grid.minor = element_blank(), # Remove grid
        panel.border = element_blank(), # Remove grid
        panel.background = element_blank()) + # Remove grid 
  scale_color_manual(values = c("Shot" = "blue3", "Goal" = "red2")) +
  scale_size_manual(values = c("Shot" = 1.5, "Goal" = 2)) +
  scale_x_continuous(limits = c(-100,100), breaks = c(-100, -50, 0, 50, 100))+
  scale_y_continuous(limits = c(-40,40), breaks = c(-40,-20,0,20,40))+
  labs(x = "Sideline", y = "Baseline",
       title = "NHL Shot Layout",
       subtitle = "Wrap-around shot type")

# Deflected
ggplot(NHL_Shot_Analysis_Data[NHL_Shot_Analysis_Data$secondaryType == "Deflected",],
       aes(x = x, y = y, group = event, color = event, size = event))+
    annotation_custom(rink, -100, 100, -40, 40) + 
  geom_point() +
  theme_bw() +
  theme(panel.grid.major = element_blank(), # Remove grid
        panel.grid.minor = element_blank(), # Remove grid
        panel.border = element_blank(), # Remove grid
        panel.background = element_blank()) + # Remove grid 
  scale_color_manual(values = c("Shot" = "blue3", "Goal" = "red2")) +
  scale_size_manual(values = c("Shot" = 1.5, "Goal" = 2)) +
  scale_x_continuous(limits = c(-100,100), breaks = c(-100, -50, 0, 50, 100))+
  scale_y_continuous(limits = c(-40,40), breaks = c(-40,-20,0,20,40))+
  labs(x = "Sideline", y = "Baseline",
       title = "NHL Shot Layout",
       subtitle = "Deflected shot type")

# Backhand
# The original size of the category is too large to present in the graph. Select a sample size of 20,000
set.seed(123)
ggplot(sample_n(NHL_Shot_Analysis_Data[NHL_Shot_Analysis_Data$secondaryType == "Backhand",],20000),
       aes(x = x, y = y, group = event, color = event, size = event))+
    annotation_custom(rink, -100, 100, -40, 40) + 
  geom_point() +
  theme_bw() +
  theme(panel.grid.major = element_blank(), # Remove grid
        panel.grid.minor = element_blank(), # Remove grid
        panel.border = element_blank(), # Remove grid
        panel.background = element_blank()) + # Remove grid 
  scale_color_manual(values = c("Shot" = "blue3", "Goal" = "red2")) +
  scale_size_manual(values = c("Shot" = 1.5, "Goal" = 2)) +
  scale_x_continuous(limits = c(-100,100), breaks = c(-100, -50, 0, 50, 100))+
  scale_y_continuous(limits = c(-40,40), breaks = c(-40,-20,0,20,40))+
  labs(x = "Sideline", y = "Baseline",
       title = "NHL Shot Layout",
       subtitle = "Backhand shot type")

# Snap Shot
set.seed(123)
ggplot(sample_n(NHL_Shot_Analysis_Data[NHL_Shot_Analysis_Data$secondaryType == "Snap Shot",],20000),
       aes(x = x, y = y, group = event, color = event, size = event))+
    annotation_custom(rink, -100, 100, -40, 40) + 
  geom_point() +
  theme_bw() +
  theme(panel.grid.major = element_blank(), # Remove grid
        panel.grid.minor = element_blank(), # Remove grid
        panel.border = element_blank(), # Remove grid
        panel.background = element_blank()) + # Remove grid 
  scale_color_manual(values = c("Shot" = "blue3", "Goal" = "red2")) +
  scale_size_manual(values = c("Shot" = 1.5, "Goal" = 2)) +
  scale_x_continuous(limits = c(-100,100), breaks = c(-100, -50, 0, 50, 100))+
  scale_y_continuous(limits = c(-40,40), breaks = c(-40,-20,0,20,40))+
  labs(x = "Sideline", y = "Baseline",
       title = "NHL Shot Layout",
       subtitle = "Snap shot type")


# Slap shot
set.seed(123)
ggplot(sample_n(NHL_Shot_Analysis_Data[NHL_Shot_Analysis_Data$secondaryType == "Slap Shot",],20000),
       aes(x = x, y = y, group = event, color = event, size = event))+
    annotation_custom(rink, -100, 100, -40, 40) + 
  geom_point() +
  theme_bw() +
  theme(panel.grid.major = element_blank(), # Remove grid
        panel.grid.minor = element_blank(), # Remove grid
        panel.border = element_blank(), # Remove grid
        panel.background = element_blank()) + # Remove grid 
  scale_color_manual(values = c("Shot" = "blue3", "Goal" = "red2")) +
  scale_size_manual(values = c("Shot" = 1.5, "Goal" = 2)) +
  scale_x_continuous(limits = c(-100,100), breaks = c(-100, -50, 0, 50, 100))+
  scale_y_continuous(limits = c(-40,40), breaks = c(-40,-20,0,20,40))+
  labs(x = "Sideline", y = "Baseline",
       title = "NHL Shot Layout",
       subtitle = "Slap shot type")

#Wrist shot
set.seed(123)
ggplot(sample_n(NHL_Shot_Analysis_Data[NHL_Shot_Analysis_Data$secondaryType == "Wrist Shot",],20000),
       aes(x = x, y = y, group = event, color = event, size = event))+
    annotation_custom(rink, -100, 100, -40, 40) + 
  geom_point() +
  theme_bw() +
  theme(panel.grid.major = element_blank(), # Remove grid
        panel.grid.minor = element_blank(), # Remove grid
        panel.border = element_blank(), # Remove grid
        panel.background = element_blank()) + # Remove grid 
  scale_color_manual(values = c("Shot" = "blue3", "Goal" = "red2")) +
  scale_size_manual(values = c("Shot" = 1.5, "Goal" = 2)) +
  scale_x_continuous(limits = c(-100,100), breaks = c(-100, -50, 0, 50, 100))+
  scale_y_continuous(limits = c(-40,40), breaks = c(-40,-20,0,20,40))+
  labs(x = "Sideline", y = "Baseline",
       title = "NHL Shot Layout",
       subtitle = "Wrist shot type")

# Tip-in
ggplot(NHL_Shot_Analysis_Data[NHL_Shot_Analysis_Data$secondaryType == "Tip-In",],
       aes(x = x, y = y, group = event, color = event, size = event))+
    annotation_custom(rink, -100, 100, -40, 40) + 
  geom_point() +
  theme_bw() +
  theme(panel.grid.major = element_blank(), # Remove grid
        panel.grid.minor = element_blank(), # Remove grid
        panel.border = element_blank(), # Remove grid
        panel.background = element_blank()) + # Remove grid 
  scale_color_manual(values = c("Shot" = "blue3", "Goal" = "red2")) +
  scale_size_manual(values = c("Shot" = 1.5, "Goal" = 2)) +
  scale_x_continuous(limits = c(-100,100), breaks = c(-100, -50, 0, 50, 100))+
  scale_y_continuous(limits = c(-40,40), breaks = c(-40,-20,0,20,40))+
  labs(x = "Sideline", y = "Baseline",
       title = "NHL Shot Layout",
       subtitle = "Tip-In shot type")
```

Goal Heatmap Analysis
```{r}
OnlyGoal <- NHL_Shot_Analysis_Data[NHL_Shot_Analysis_Data$event == "Goal",]

# Wrap-around
ggplot(OnlyGoal[OnlyGoal$secondaryType == "Wrap-around",],
       aes(x = x, y = y))+
    annotation_custom(rink, -100, 100, -40, 40) + 
    geom_bin2d( bins = 30)+
    theme_bw() +
    theme(panel.grid.major = element_blank(), # Remove grid
      panel.grid.minor = element_blank(), # Remove grid
      panel.border = element_blank(), # Remove grid
      panel.background = element_blank()) + # Remove grid 
    scale_x_continuous(limits = c(-100,100), breaks = c(-100, -50, 0, 50, 100))+
    scale_y_continuous(limits = c(-40,40), breaks = c(-40,-20,0,20,40))+
      labs(x = "Sideline", y = "Baseline",
      title = "NHL Goal Density Heatmap",
      subtitle = "Wrap-around shot type")

# Deflected
ggplot(OnlyGoal[OnlyGoal$secondaryType == "Deflected",],
       aes(x = x, y = y))+
    annotation_custom(rink, -100, 100, -40, 40) + 
    geom_bin2d()+
    theme_bw() +
    theme(panel.grid.major = element_blank(), # Remove grid
      panel.grid.minor = element_blank(), # Remove grid
      panel.border = element_blank(), # Remove grid
      panel.background = element_blank()) + # Remove grid 
    scale_x_continuous(limits = c(-100,100), breaks = c(-100, -50, 0, 50, 100))+
    scale_y_continuous(limits = c(-40,40), breaks = c(-40,-20,0,20,40))+
      labs(x = "Sideline", y = "Baseline",
      title = "NHL Goal Density Heatmap",
      subtitle = "Deflected shot type")

# Backhand
ggplot(OnlyGoal[OnlyGoal$secondaryType == "Backhand",],
       aes(x = x, y = y))+
    annotation_custom(rink, -100, 100, -40, 40) + 
    geom_bin2d()+
    theme_bw() +
    theme(panel.grid.major = element_blank(), # Remove grid
      panel.grid.minor = element_blank(), # Remove grid
      panel.border = element_blank(), # Remove grid
      panel.background = element_blank()) + # Remove grid 
    scale_x_continuous(limits = c(-100,100), breaks = c(-100, -50, 0, 50, 100))+
    scale_y_continuous(limits = c(-40,40), breaks = c(-40,-20,0,20,40))+
      labs(x = "Sideline", y = "Baseline",
      title = "NHL Goal Density Heatmap",
      subtitle = "Backhand shot type")

# Snap Shot
ggplot(OnlyGoal[OnlyGoal$secondaryType == "Snap Shot",],
       aes(x = x, y = y))+
    annotation_custom(rink, -100, 100, -40, 40) + 
    geom_bin2d()+
    theme_bw() +
    theme(panel.grid.major = element_blank(), # Remove grid
      panel.grid.minor = element_blank(), # Remove grid
      panel.border = element_blank(), # Remove grid
      panel.background = element_blank()) + # Remove grid 
    scale_x_continuous(limits = c(-100,100), breaks = c(-100, -50, 0, 50, 100))+
    scale_y_continuous(limits = c(-40,40), breaks = c(-40,-20,0,20,40))+
      labs(x = "Sideline", y = "Baseline",
      title = "NHL Goal Density Heatmap",
      subtitle = "Snap Shot shot type")

# Slap Shot
ggplot(OnlyGoal[OnlyGoal$secondaryType == "Slap Shot",],
       aes(x = x, y = y))+
    annotation_custom(rink, -100, 100, -40, 40) + 
    geom_bin2d()+
    theme_bw() +
    theme(panel.grid.major = element_blank(), # Remove grid
      panel.grid.minor = element_blank(), # Remove grid
      panel.border = element_blank(), # Remove grid
      panel.background = element_blank()) + # Remove grid 
    scale_x_continuous(limits = c(-100,100), breaks = c(-100, -50, 0, 50, 100))+
    scale_y_continuous(limits = c(-40,40), breaks = c(-40,-20,0,20,40))+
      labs(x = "Sideline", y = "Baseline",
      title = "NHL Goal Density Heatmap",
      subtitle = "Slap Shot shot type")

#Wrist Shot
set.seed(123)
ggplot(sample_n(OnlyGoal[OnlyGoal$secondaryType == "Wrist Shot",],10000),
       aes(x = x, y = y))+
    annotation_custom(rink, -100, 100, -40, 40) + 
    geom_bin2d()+
    theme_bw() +
    theme(panel.grid.major = element_blank(), # Remove grid
      panel.grid.minor = element_blank(), # Remove grid
      panel.border = element_blank(), # Remove grid
      panel.background = element_blank()) + # Remove grid 
    scale_x_continuous(limits = c(-100,100), breaks = c(-100, -50, 0, 50, 100))+
    scale_y_continuous(limits = c(-40,40), breaks = c(-40,-20,0,20,40))+
      labs(x = "Sideline", y = "Baseline",
      title = "NHL Goal Density Heatmap",
      subtitle = "Wrist Shot shot type")

# Tip-in
ggplot(OnlyGoal[OnlyGoal$secondaryType == "Tip-In",],
       aes(x = x, y = y))+
    annotation_custom(rink, -100, 100, -40, 40) + 
    geom_bin2d()+
    theme_bw() +
    theme(panel.grid.major = element_blank(), # Remove grid
      panel.grid.minor = element_blank(), # Remove grid
      panel.border = element_blank(), # Remove grid
      panel.background = element_blank()) + # Remove grid 
    scale_x_continuous(limits = c(-100,100), breaks = c(-100, -50, 0, 50, 100))+
    scale_y_continuous(limits = c(-40,40), breaks = c(-40,-20,0,20,40))+
      labs(x = "Sideline", y = "Baseline",
      title = "NHL Goal Density Heatmap",
      subtitle = "Tip-In shot type")
```
